---
title: "Course"
author: "Shashwat"
date: "January 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Background

Subjects were asked to perform barbell lifts correctly and incorrectly in 5 different ways.
* Exactly according to the specification (Class A)
* Throwing the elbows to the front (Class B) - mistake
* Lifting the dumbbell only halfway (Class C) - mistake
* Lowering the dumbbell only halfway (Class D) - mistake
* Throwing the hips to the front (Class E) - mistake

Accelerometers were located on
1. belt
2. forearm
3. arm
Task

Create a report describing
* how you built your model,
* how you used cross validation
* what you think the expected out of sample error is
* why you made the choices you did

###Loading libraries
```{r}
#install.packages("doParallel")
#install.packages("randomForest")
#install.packages("e1071")
suppressWarnings(suppressMessages(library(caret)))
suppressWarnings(suppressMessages(library(randomForest)))
suppressWarnings(suppressMessages(library(e1071)))
set.seed(1603)
```

###QUESTION

Create a model to predict the manner in which the subjects did the exercise using the accelerometer data as predictors.
The outcome to be predicted is the "classe" variable.

###Download data.
```{r}
trainingFilename   <- 'pml-training.csv'
quizFilename       <- 'pml-testing.csv'
trainingUrl        <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'
quizUrl            <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'
```

###Cleaning data.
```{r}
training.df     <-read.csv(trainingFilename, na.strings=c("NA","","#DIV/0!"))
training.df     <-training.df[,colSums(is.na(training.df)) == 0]
dim(training.df) #;head(training.df,3)
```

```{r}
quiz.df         <-read.csv(quizFilename , na.strings=c("NA", "", "#DIV/0!"))
quiz.df         <-quiz.df[,colSums(is.na(quiz.df)) == 0]
dim(quiz.df)
```

###Feature selection.
```{r}
Training.df   <-training.df[,-c(1:7)]
Quiz.df <-quiz.df[,-c(1:7)]
dim(Training.df)
```

Non zero values in data
```{r}
Training.nzv<-nzv(Training.df[,-ncol(Training.df)],saveMetrics=TRUE)
```
Displaying names
```{r}
rownames(Training.nzv)
```
Checking dimensions.
```{r}
dim(Training.nzv)[1]
```

###Data partition
```{r}
inTrain     <- createDataPartition(Training.df$classe, p = 0.6, list = FALSE)
inTraining  <- Training.df[inTrain,]
inTest      <- Training.df[-inTrain,]
dim(inTraining);dim(inTest)
```

###Constructing the model using cross validation
```{r}
myModelFilename <- "myModel.RData"
if (!file.exists(myModelFilename)) {

    # Parallel cores  
    #require(parallel)
    library(doParallel)
    ncores <- makeCluster(detectCores() - 1)
    registerDoParallel(cores=ncores)
    getDoParWorkers() # 3    
    
    # use Random Forest method with Cross Validation, 4 folds
    myModel <- train(classe ~ .
                , data = inTraining
                , method = "rf"
                , metric = "Accuracy"  # categorical outcome variable so choose accuracy
                , preProcess=c("center", "scale") # attempt to improve accuracy by normalising
                , trControl=trainControl(method = "cv"
                                        , number = 4 # folds of the training data
                                        , p= 0.60
                                        , allowParallel = TRUE 
#                                       , seeds=NA # don't let workers set seed 
                                        )
                )

    save(myModel, file = "myModel.RData")
    # 3:42 .. 3:49 without preProcess
    # 3:51 .. 3:58 with preProcess
    stopCluster(ncores)
} else {
    # Use cached model  
    load(file = myModelFilename, verbose = TRUE)
}
```

Printing the model
```{r}
print(myModel, digits=4)
```

###Prediction.
```{r}
predTest <- predict(myModel, newdata=inTest)
```

###Evalution
####Confusion matrix.
```{r}
confusionMatrix(predTest, inTest$classe)
```

Out of sample error-
The out-of-sample error of 0.0037 or 0.37%.

Accuracy is very high, at 0.9963, and this figure lies within the 95% confidence interval.

###Final model.
```{r}
myModel$finalModel
```
```{r}
varImp(myModel)
```
###Conclusion.
```{r}
print(predict(myModel, newdata=Quiz.df))
```